---
title: "Plotting MVBS in R"
author: "Erin LaBrecque"
date: '2025-04-05'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`tidync` was developed and maintained by [ROpenSci](https://github.com/ropensci/tidync/tree/main)

# Install... 

from CRAN:    
```{r}
#install.packages("tidync")
```

or the developmental install from GitHub:  
```{r}
# install.packages("remotes")
#remotes::install_github("ropensci/tidync", dependencies = TRUE)
```

# Attach packages  

```{r message=FALSE}
library(tidync)
library(dplyr)
library(ggplot2)
library(lubridate)
```


# Load the data

The .nc file is in the data folder. First set a path to the data folder using the `here` package, then load the file using `tidync`.

```{r}
# path to data folder
dta <- here::here("data")
```

We are only going to use one function from the `here` package, so we are not going to attach it to session like we did by calling the other packages with `library()`. By using the package name and `::` we can call a function from the package (assuming it is in your library folder). We are using `here()` instead of `setwd()` because `here()` enables easy file referencing by using the your top-level directory (folder) to build file paths. Use `paste0()` to concatenate the file path and file name.

paste0(dta, "/test_MVBS.nc") -> "path to your folder/data/test_MVBS.nc"


```{r}
# load data - retrieve both values in the data array and the metadata with tidync()
mvbs <- tidync(paste0(dta, "/test_MVBS.nc"))
```


# Examine the data structure
```{r}
# show data
mvbs
```

Under dimensions, `unlim == FALSE` means that the extent of the data are fixed. `coord_dim == TRUE` means that the dimension being used is attached o the dataset. In larger datasets, you might not want to bring in all the data.  

Use `hyper_dims()` to inspect the dimensions in more detail.

```{r}
hyper_dims(mvbs)
```

<br>

Use `hyper_vars()` to explore the variable - Sv.

```{r}
hyper_vars(mvbs)
```

# Examine the data values

To access the data values, you have to active the correct grid with `activate()`, then extract the values with `hyper_tibble()`.

Use `glimpse()` to view the data types and the first few values.
```{r}
ping_data <- mvbs %>% 
  activate(Sv) %>% 
  hyper_tibble()

glimpse(ping_data)
```
 
 <br> 
 
Sv is stored as double, but depth and ping_time are stored as characters. These need to be changed to double and datetime respectively.

```{r}
ping_data <- ping_data %>%  
  mutate(depth = as.numeric(depth),
         ping_time = ymd_hms(ping_time))

glimpse(ping_data)
```

<br> 

# Plot 

```{r}
ping_data %>% 
  ggplot(aes(ping_time, depth, color = Sv)) +
  geom_tile() +
  scale_y_reverse() +
  scale_color_viridis_c(limits = c(-100, -20)) +
  facet_wrap(~channel, nrow = 3) +
  theme_minimal() +
  theme(legend.title = element_text(angle = 90, hjust = 0.5),
        legend.title.position = "right") +
  labs(x = "Ping time",
       y = "Range distance (m)",
       color = "Mean volume backscatteing strength
       (MVBS, mean Sv re 1 m-1) [dB]")
```

