---
title: "Plotting MVBS in R"
author: "Erin LaBrecque"
date: '2025-04-05'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`tidync` was developed and maintained by [ROpenSci](https://github.com/ropensci/tidync/tree/main)

# Install the package... 

from CRAN:    
```{r}
#install.packages("tidync")
```

or the developmental install from GitHub:  
```{r}
# install.packages("remotes")
#remotes::install_github("ropensci/tidync", dependencies = TRUE)
```

# Attach packages  

The warnings have been muted with `message = FALSE`
```{r message=FALSE}
library(tidync)
library(dplyr)
library(ggplot2)
library(lubridate)
```


# Loading the data

The .nc file is in the data folder. First set a path to the data folder using the `here` package, then load the file using `tidync`.

We are only going to use one function from the `here` package, so we are not going to attach it to session like we did by calling the other packages with `library()`. By using the package name and `::` we can call a function from the package. We are using `here()` instead of `setwd()` because `here` enables easy file referencing by using the your top-level directory (folder) to build file paths.  


```{r}
# path to data folder
dta <- here::here("data")

# load data
mvbs <- tidync(paste0(dta, "/test_MVBS.nc"))

# show data
mvbs
```

There are three dimensions in the file. Use `hyper_dims` to inspect the dimensions in more detail.

```{r}
hyper_dims(mvbs)
```

<br>


```{r}
hyper_vars(mvbs)
```

